Modal Verification Patterns for Systems

### 2025-06-23 paper.tex
+

\todo[inline,color=cyan]{Ismail, is this text copy-pasted from the VMM paper? 
It has phrasing like ``In this paper...''
That's actually a major issue, self-plagiarism of another peer-reviewed piece of work can get
you into a lot of trouble.
}\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
\paragraph{Virtual Memory Managers with Multiple Address-Spaces}
\label{sec:backgroundonvmm}
\todo{change the text}
An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses. A typical OS manages multiple processes at once . Each process is identified by an \emph{unique root address} of its page-table tree. As part of switching to a new process, i.e., loading the address-space mappings as the current view of the memory, the CPU stores the unique root address of the page-table tree in a specific register (\texttt{cr3}) as shown in Figure \ref{fig:pagetables}. Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes. Virtual memory managers are the modules in OS kernels handling this kind of tasks. The treat address-spaces as containers of virtual addresses so that they can bookkeep which which virtual addresses are valid (and in what way) in which address spaces.
\paragraph{Virtualization of Disk-Blocks in Filesystems}
\textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
%\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode. \textsf{ModFS}'s code is translated to \textsf{Roqc} using the \textsf{Goose}~\cite{chajed2019verifying,chajed2021gojournal} tool, which translates a significant subset of \textsf{Go} to \textsf{Roqc} to permit formal verification of \textsc{Go} code using \textsf{Iris}~\cite{jung2018iris}. \textsf{Goose} generates code in an extension of \textsf{Iris}'s \textsf{HeapLang} with libraries for disk manipulation and \textsc{Go} primitive types (e.g., maps) as shown in Figure \ref{fig:modfslayout}.

>>>>

\todo[inline,color=cyan]{Ismail, is this text copy-pasted from the VMM paper? 
It has phrasing like ``In this paper...''
That's actually a major issue, self-plagiarism of another peer-reviewed piece of work can get
you into a lot of trouble.
}\todo{inline,color=pink}{I am well-aware what counts as as (including self-plagiarism). These were annotated with todos including cautions for changing the text.}
\paragraph{Virtual Memory Managers(VMM) with Multiple Address-Spaces}
\label{sec:backgroundonvmm}
\todo{change the text}
An intuition for the abstraction that \emph{contains} a page-table tree is called an \emph{address-space} of a process whose virtual addresses to physical addresses. A typical OS manages multiple processes at once . Each process is identified by an \emph{unique root address} of its page-table tree. As part of switching to a new process, i.e., loading the address-space mappings as the current view of the memory, the CPU stores the unique root address of the page-table tree in a specific register (\texttt{cr3}) as shown in Figure \ref{fig:pagetables}. Using different mappings, which map only disjoint portions of physical memory (with some exceptions in the next section) is how the OS ensures memory isolation between processes. Virtual memory managers are the modules in OS kernels handling this kind of tasks. The treat address-spaces as containers of virtual addresses so that they can bookkeep which which virtual addresses are valid (and in what way) in which address spaces.
\paragraph{Virtualization of Disk-Blocks in Filesystems}
Another important computer resource computer resource that is \emph{virtualized} for access is disk blocks. Filesystems are computer programs that serve to other programs for their disk resource need. Like VMMs using page-table trees to fabricate more memory addresses than what actually exists, filesystems uses indexing maps for fabricating more \emph{virtual disk-page} addresses which are conventionally called \emph{disk-pages} to be distinguished from physical \emph{disk-block} addresses. 
\textsf{ModFS} is a copy-on-write (\textsf{CoW}) filesystem, which as described in the introduction performs updates by copying modified data to \emph{new and unused} locations on disk rather than updating in place, ensuring that in the absence of physical storage corruption, there is \emph{always} a fully consistent filesystem on disk, removing the need for recovery from journals in the case of a crash, which can be very expensive for large volumes.
%\textsf{ModFS} is implemented in \textsc{Go}~\cite{donovan2015go}, using the FUSE~\cite{fuse} interface to run a filesystem in usermode. \textsf{ModFS}'s code is translated to \textsf{Roqc} using the \textsf{Goose}~\cite{chajed2019verifying,chajed2021gojournal} tool, which translates a significant subset of \textsf{Go} to \textsf{Roqc} to permit formal verification of \textsc{Go} code using \textsf{Iris}~\cite{jung2018iris}. \textsf{Goose} generates code in an extension of \textsf{Iris}'s \textsf{HeapLang} with libraries for disk manipulation and \textsc{Go} primitive types (e.g., maps) as shown in Figure \ref{fig:modfslayout}.
